Predict Disease Outcome by Community Structures

### 2016-08-17 code/HumanCdGF_microbiota_predict.Rmd
add rf of all cfu by day together

@@

## Does Day Community structure predict C. difficile CFU of that day?

RF all mice/days together
RF cfu vs community by day

@@

@@

ggdraw()+
  draw_plot(day_persist_rf_plot, 0, 0.5, 0.75, 0.5) + 
  draw_plot(day_persist_rsq_plot, 0.875, 0.75, 0.125, 0.25) +
  draw_plot(day_persist_mse_plot, 0.75, 0.75, 0.125, 0.25) + 
  draw_plot(day_persist_r2feat_plot, 0.75, 0.5, 0.25, 0.25) + 
  draw_plot(day_persist_feat_plot, 0, 0, 1, 0.5)

>>>>

## Can Day 0 Community members predict C. difficile CFU?
```{r fig.height= 10, fig.width=10}

@@

## Does Community structure predict C. difficile CFU of same day?

### RF cfu vs community by day

@@

### RF all mice/days together

```{r }
predict_cfu_all_df <- CFU_community_by_day[!is.na(CFU_community_by_day$log_cfu),
                                           ! colnames(CFU_community_by_day) %in% c('Row.names','day')]
RF_cfu_all_rf <- randomForest(log_cfu ~ ., data = predict_cfu_all_df,
                   importance = TRUE, ntree = n_trees)
RF_cfu_all_df <- data.frame(cbind(CFU_community_by_day[!is.na(CFU_community_by_day$log_cfu), 
                                                       c('day','log_cfu')], 
                                  predicted = RF_cfu_all_rf$predicted))

mse_all <- tail(RF_cfu_all_rf$mse, 1)
rsq_all <- tail(RF_cfu_all_rf$rsq, 1)

all_rf_plot <- ggplot(data = RF_cfu_all_df, aes(x = log_cfu, y = predicted)) + geom_point() +
        labs(x= 'Observed CFU (log10)', y = 'Predicted CFU (log10)') + theme_bw() + 
        ggtitle("Random Forest Prediction of C Diff CFU") +
        annotate('text', x=0.15*(max(RF_cfu_all_df$log_cfu)), y=0.95*(max(RF_cfu_all_df$predicted)), 
                 label=paste('R^2 = ', tail(RF_cfu_all_rf$rsq, 1),'\nMSE = ', tail(RF_cfu_all_rf$mse, 1), sep = ''))

importance_all <- sort(importance(RF_cfu_all_rf)[,1], decreasing = T)

rsq_features_all <- data.frame('Features'= NA, 'RSQ'=NA, 'MSE'=NA)
for (k in 1:25){
    rf_partial <- randomForest(log_cfu ~ ., 
                               data=predict_cfu_all_df[ , c( 'log_cfu', names(importance_all)[1:k])], 
                               importance=TRUE, ntree=n_trees)
    rsq_features_all <- rbind(rsq_features_all, c(k, tail(rf_partial$rsq, 1), tail(rf_partial$mse, 1)))
}
rsq_features_all <- rsq_features_all[-1,]
all_r2feat_plot <- ggplot(rsq_features_all, aes(x=Features, y=RSQ)) + geom_line() +
  labs(x = 'Features (n)', y = 'R^2') + scale_color_discrete(guide = FALSE) + theme_bw()

important_OTU_all <- data.frame(head(importance_all, n_important))
colnames(important_OTU_all) <- 'Importance'
important_OTU_all$OTU <- rownames(important_OTU_all)

all_feat_plot <- ggplot(data = important_OTU_all, aes(x = factor(OTU), y = Importance)) + 
        geom_point() + scale_x_discrete(limits = rev(important_OTU_all$OTU)) +
        labs(x= '', y = '% Increase in MSE') + theme_bw() + coord_flip()

ggdraw()+
  draw_plot(all_rf_plot, 0, 0.5, 0.5, 0.5) + 
  draw_plot(all_r2feat_plot, 0, 0, 0.5, 0.5) +
  draw_plot(all_feat_plot, 0.5, 0, 0.5, 1)

```{r fig.width=10, fig.height=5}
predict_cfu_all_df <- CFU_community_by_day[!is.na(CFU_community_by_day$log_cfu),
                                           ! colnames(CFU_community_by_day) %in% c('Row.names','day')]
RF_cfu_all_rf <- randomForest(log_cfu ~ ., data = predict_cfu_all_df,
                   importance = TRUE, ntree = n_trees)
RF_cfu_all_df <- data.frame(cbind(CFU_community_by_day[!is.na(CFU_community_by_day$log_cfu), 
                                                       c('day','log_cfu')], 
                                  predicted = RF_cfu_all_rf$predicted))

mse_all <- tail(RF_cfu_all_rf$mse, 1)
rsq_all <- tail(RF_cfu_all_rf$rsq, 1)

all_rf_plot <- ggplot(data = RF_cfu_all_df, aes(x = log_cfu, y = predicted)) + geom_point() +
        labs(x= 'Observed CFU (log10)', y = 'Predicted CFU (log10)') + theme_bw() + 
        ggtitle("Random Forest Prediction of C Diff CFU") +
        annotate('text', x=0.15*(max(RF_cfu_all_df$log_cfu)), y=0.95*(max(RF_cfu_all_df$predicted)), 
                 label=paste('R^2 = ', tail(RF_cfu_all_rf$rsq, 1),'\nMSE = ', tail(RF_cfu_all_rf$mse, 1), sep = ''))

importance_all <- sort(importance(RF_cfu_all_rf)[,1], decreasing = T)

rsq_features_all <- data.frame('Features'= NA, 'RSQ'=NA, 'MSE'=NA)
for (k in 1:25){
    rf_partial <- randomForest(log_cfu ~ ., 
                               data=predict_cfu_all_df[ , c( 'log_cfu', names(importance_all)[1:k])], 
                               importance=TRUE, ntree=n_trees)
    rsq_features_all <- rbind(rsq_features_all, c(k, tail(rf_partial$rsq, 1), tail(rf_partial$mse, 1)))
}
rsq_features_all <- rsq_features_all[-1,]
all_r2feat_plot <- ggplot(rsq_features_all, aes(x=Features, y=RSQ)) + geom_line() +
  labs(x = 'Features (n)', y = 'R^2') + scale_color_discrete(guide = FALSE) + theme_bw()

important_OTU_all <- data.frame(head(importance_all, n_important))
colnames(important_OTU_all) <- 'Importance'
important_OTU_all$OTU <- rownames(important_OTU_all)

all_feat_plot <- ggplot(data = important_OTU_all, aes(x = factor(OTU), y = Importance)) + 
        geom_point() + scale_x_discrete(limits = rev(important_OTU_all$OTU)) +
        labs(x= '', y = '% Increase in MSE') + theme_bw() + coord_flip()

ggdraw()+
  draw_plot(all_rf_plot, 0, 0.5, 0.5, 0.5) + 
  draw_plot(all_r2feat_plot, 0, 0, 0.5, 0.5) +
  draw_plot(all_feat_plot, 0.5, 0, 0.5, 1)

@@

ggdraw()+
  draw_plot(day_persist_rf_plot, 0, 0.5, 0.75, 0.5) + 
  draw_plot(day_persist_rsq_plot, 0.875, 0.75, 0.125, 0.25) +
  draw_plot(day_persist_mse_plot, 0.75, 0.75, 0.125, 0.25) + 
  draw_plot(day_persist_r2feat_plot, 0.75, 0.5, 0.25, 0.25) + 
  draw_plot(day_persist_feat_plot, 0, 0, 1, 0.5)
```

#### All CFU compared to same day community without mice euthanized early 
```{r fig.width=10, fig.height=5}

predict_cfu_persist_df <- without_by_day[!is.na(without_by_day$log_cfu),
                                           ! colnames(without_by_day) %in% c('Row.names','day','Mouse_ID')]
RF_cfu_persist_rf <- randomForest(log_cfu ~ ., data = predict_cfu_persist_df,
                   importance = TRUE, ntree = n_trees)
RF_cfu_persist_df <- data.frame(cbind(without_by_day[!is.na(without_by_day$log_cfu), 
                                                       c('day','log_cfu')], 
                                  predicted = RF_cfu_persist_rf$predicted))

mse_all <- tail(RF_cfu_persist_rf$mse, 1)
rsq_all <- tail(RF_cfu_persist_rf$rsq, 1)

all_persist_rf_plot <- ggplot(data = RF_cfu_persist_df, aes(x = log_cfu, y = predicted)) + geom_point() +
        labs(x= 'Observed CFU (log10)', y = 'Predicted CFU (log10)') + theme_bw() + 
        ggtitle("Random Forest Prediction of C Diff CFU") +
        annotate('text', x=0.15*(max(RF_cfu_persist_df$log_cfu)), y=0.95*(max(RF_cfu_persist_df$predicted)), 
                 label=paste('R^2 = ', tail(RF_cfu_persist_rf$rsq, 1),'\nMSE = ', tail(RF_cfu_persist_rf$mse, 1), sep = ''))

importance_all_persist <- sort(importance(RF_cfu_persist_rf)[,1], decreasing = T)

rsq_features_all_persist <- data.frame('Features'= NA, 'RSQ'=NA, 'MSE'=NA)
for (k in 1:25){
    rf_partial <- randomForest(log_cfu ~ ., 
                               data=predict_cfu_persist_df[ , c( 'log_cfu', names(importance_all_persist)[1:k])], 
                               importance=TRUE, ntree=n_trees)
    rsq_features_all_persist <- rbind(rsq_features_all_persist, c(k, tail(rf_partial$rsq, 1), tail(rf_partial$mse, 1)))
}
rsq_features_all_persist <- rsq_features_all_persist[-1,]
all_persist_r2feat_plot <- ggplot(rsq_features_all_persist, aes(x=Features, y=RSQ)) + geom_line() +
  labs(x = 'Features (n)', y = 'R^2') + scale_color_discrete(guide = FALSE) + theme_bw()

important_OTU_all_persist <- data.frame(head(importance_all_persist, n_important))
colnames(important_OTU_all_persist) <- 'Importance'
important_OTU_all_persist$OTU <- rownames(important_OTU_all_persist)

all_persist_feat_plot <- ggplot(data = important_OTU_all_persist, aes(x = factor(OTU), y = Importance)) + 
        geom_point() + scale_x_discrete(limits = rev(important_OTU_all_persist$OTU)) +
        labs(x= '', y = '% Increase in MSE') + theme_bw() + coord_flip()

ggdraw()+
  draw_plot(all_persist_rf_plot, 0, 0.5, 0.5, 0.5) + 
  draw_plot(all_persist_r2feat_plot, 0, 0, 0.5, 0.5) +
  draw_plot(all_persist_feat_plot, 0.5, 0, 0.5, 1)
